ediblecity: an R package to model and estimate the benefits of urban agriculture

### 2023-05-09 paper_v01.Rmd
version to submit

---
title: "ediblecity: an R package to model and estimate the benefits of urban agriculture"
author:
- Josep Pueyo-Ros \textsuperscript{1,2,*}
- Joaquim Comas \textsuperscript{1,3}
- Lluís Corominas \textsuperscript{1,2}
output: 
  word_document:
    # number_sections: true
    # keep_tex: true
bibliography: library.bib
---

@@

Urban agriculture is gaining attraction to become one of the pillars of the urban ecological transition and to increase food security in an urbanized planet. However, there is a lack of systematic quantification of the benefits provided by urban agriculture solutions. In this paper, we present an R package to estimate several indicators related to benefits of urban agriculture. The goal is to provide a tool for researchers and practitioners interested in the impacts of urban agriculture. The ediblecity package provides functions to calculate 8 indicators: urban heat island, runoff prevention, green areas accessibility, NO~2~ sequestration, jobs created in commercial gardens, volunteers involved in community gardens, green per capita and, finally, food production. Moreover, the package also provides a function to create scenarios with different implementations of urban agriculture. We illustrate the use of the package by comparing three scenarios in a neighborhood of Girona (Spain), which is included in the package as an example dataset. There, we compare scenarios with an increasing amount of urban agriculture solutions. The `ediblecity` package is open-source software. This allows other R developers to contribute to the package providing new functionalities or improving the existing ones.

@@

Therefore, decision-makers, who are responsible of leading the urban transitions to more sustainable and resilient cities, are orphan of evidence in how to implement urban agriculture to maximize its impact on sustainability. Yet, several studies provided some insights that can guide the implementation of urban agriculture. For instance, some models assist planners in rainwater harvesting [@Gittleman2017; @Lupia2017]. Broader, @GomezVillarino2021 developed guidelines to maximize ecosystem services through urban agriculture. And, as expected, many models have been developed to quantify food production and food security provided by urban agriculture [@Grafius2020; @Grewal2012; @Hsieh2017; @MacRae2010].

Hence, our goal is to provide a tool to estimate multiple benefits of urban agriculture that help decision-makers to strategically implement urban agriculture solutions. We developed the estimations for eight indicators and a functionality to create scenarios of urban agriculture based on the proportion of elements to be transformed to urban agriculture and on which elements will be transformed and packed all of them in an R package called `ediblecity`. In the methods, we present the main functionalities of the package and then we apply those to an example in a neighborhood of Girona (Spain).

@@

## Implementation

@@

The `ediblecity` package provides 8 functions to estimate 8 different indicators and a function to create scenarios. It takes as a basis a spatial representation of a city (or a part of a city) as a GIS layer. This representation must have one attribute indicating the land uses of the city (aka functions), such as gardens, streets, rooftops, etc. Some indicators require specific information to be included in the representation. These is explained in each indicator's section.

```{r}
color_names <- sort(unique(city_example$Function))
colors <- c("Brown", "Lightgreen", "chartreuse4", "Grey20", 
            "Black", "chocolate2", "Yellow", "Lightblue",
            "Red", "Grey50", "Darkgreen", "Yellow"
            )
names(colors) <- color_names

@@

city_example |> 
  ggplot()+
  geom_sf(aes(fill=Function), color="white", linewidth = 0.1)+
  scale_fill_manual(values = colors)+
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        axis.text = element_text(size=6))

@@

tribble(
  ~Column, ~Description,
  "Function", "A category representing the urban elements",
  "Function_verbose", "A more detailed category for the element, for example, if it is residential",
  "floors", "The number of floors of the element, 0 for non-built",
  "area", "The surface of the element",
  "flat_area", "The surface of the element that is flat (slope < 5º)",
  "edible_area", "The surface that is used to grow edible plants. Only applicable to urban agriculture solutions"
) |> 
  kable(caption = "Structure of the urban representation example")

@@

Along with the example for an urban representation, the `ediblecity` package also includes a `data.frame` with the general attributes of each green typology in the urban representation used to estimate the indicators (Table 2). However, the user can provide their own attributes to estimate any indicator. In `city_functions` there are other columns not shown in table 2, they are logical variables (i.e. TRUE/FALSE) used internally by the package to select urban agriculture elements.

`pGreen` is the proportion of green of the urban element. In urban agriculture solutions, this is overriden by the attribute `edible_area`. The following attributes come in pairs. This is to consider uncertainty in the estimations. The functions use a random value within the range provided by the pair of values for each element in the city. `no2_seq` is the capacity of the element to capture NO~2~ in gr/s. `food` is the food productivity in kg/m^2^ and `CN` is their curve number, used to calculate infiltration rates [@Cronshey1985].

@@

city_functions |>
  select(-starts_with("...")) |> 
  select(-edible, -public, -jobs, -volunteers, -location) |>
  select(-starts_with("water")) |> 
  kable(caption = "General attributes of the elements of urban green used to estimate the indicators")
```
### Indicators estimated

@@

The urban heat island is a measure of how urban agriculture can contribute to climate change adaptation [@Langemeyer2021]. The indicator uses the equation developed by @Theeuwes2017, which was validated in 14 cities. It calculates the difference in air temperature between the urban street canyon and the rural environment. To calculate this indicator, the user must provide a raster representing the sky view factor (SVF), which describes the proportion of the unobstructed hemisphere above a certain location. SAGA, a collection of open-source algorithms for geocomputation, provides an algorithm to calculate the SVF [@Conrad2008].

@@

The indicator to estimate the urban heat island is implemented by the package under the function `UHI`. The user must provide the urban representation (`x`) and the raster with SVF values (`SVF`). The `green_df` argument is a `data.frame` with the proportion of green of each element (`Function`) in the urban representation. All the meteorological arguments are provided by default, based on the example provided (Mediterranean climate). However, the user can override them to provide values of their city of interest.

@@

UHI(
  x,
  SVF,
  green_df = NULL,
  Qql = 6.11,
  Cair = 1007,
  Pair = 1.14,
  Tmax = 30.8,
  Tmin = 20,
  windspeed = 2.77,
  return_raster = FALSE,
  verbose = FALSE
)

@@

where $P$ is the rainfall volume in mm; $I_a$ is the initial abstraction (all losses before runoff begins); and $S$ is the potential soil moisture retention, which is a function of the curve number (which is detemined by hydrologic soil group, see @Cronshey1985 for more details on the method). The SCS generalizes $I_a$ as $0.2S$, we modified this generalization to include the rainwater harvested:

@@

The `runoff_prev` function estimates the runoff (in mm) as well as the total rainfall in `x` and the total rainwater harvested. Along with the urban representation (`x`), the user must provide a `data.frame` with the functions, two columns representing the range of curve numbers and a logical column indicating if the element (a urban garden, a building,...) has potential to harvest rainwater. The argument `rain` allows to set the rain event (in mm), which must be defined by the user. The curve number of each element is randomized within this range provided in `runoff_df`. If `runoff_df` is not provided, `city_functions` is used instead. Following, `floors_field` is the name of attribute in `x` that specifies the number of floors of each element; `harvest_dist` is the maximum distance to consider that a building is adjacent to the element; and `tank_size` is a range for the volume of the rainwater tank (in m^3^). The volume of each tank in the city is randomized within this range. All randomizations follow a random uniform distribution within the correspondent range.

@@

runoff_prev(
  x,
  runoff_df = NULL,
  rain = 85,
  floors_field = "floors",
  harvest_dist = 10,
  tank_size = c(0, 45)
)
```

@@

This indicator calculates the distance from each home to the closest public green area and return a summary of statistics (min, 25%, 50%, mean, 75% and max). It includes the possibility to exclude areas smaller than a threshold. The function to calculate these distances is `green_distance`. It requires, as usual, the urban representation (`x`) and a vector with all the categories considered public green areas. If it is not provided, the function uses the categories from `city_functions` where the attribute `public` is `TRUE`. The argument `residence_col` indicates the variable of the urban representation (`x`) that must be used to identify the residences. Subsequently, `residences` indicates which categories of `residence_col` must be considered.  The `min_area` argument can be used to exclude smaller areas than the value passed to the function (the threshold mentioned above).

@@

Finally, if `verbose` argument is set to `TRUE`, a vector with all distances is returned.

@@

green_distance(
  x,
  green_cat = NULL,
  residence_col = "Function_verbose",
  residences = "Residence",
  min_area = 5000,
  percent_out = FALSE,
  max_dist = 300,
  verbose = FALSE
)

@@

$$
seq = \sum_{i=1}^{N}{a_i \times cap_i} \times 1000
$$

where $a_i$ is the area (in m^2^) of the element $i$; and $cap_i$ is the capacity of element $i$ to sequester NO~2~ (in $\mu$g/s).

The function to estimate the NO~2~ sequestered is `no2_seq`. It has only two arguments; the urban representation (`x`) and a `data.frame` with four columns:

- `functions`: Column with the function to be considered in the calculations corresponding
 to `Function` attribute in `x`.
- `no2_seq1`: The low range of NO2 sequestration of each function (in $\mu$g/s/m^2^).
- `no2_seq2`: The high range of NO2 sequestration of each function (in $\mu$g/s/m^2^).
- `pGreen`: The proportion of green surface in each function.

@@

no2_seq(x, green_df = NULL)

@@

Two indicators are proposed to account for the hours of work in the ECS. One indicator relates to volunteers time and the other one relates to time of new workers (jobs created). When the urban agriculture solutions are community solutions they need volunteers to be involved. On the other hand, when they are for commercial purposes, they are supposed to create jobs. Both indicators use the same equation:

@@

where $a_i$ is the area in m^2^ used to grow plants (`edible_area`) in the element $i$; and $k$ is the number of jobs or volunteers by m^2^. In both functions, $k$ is sampled from a random uniform distribution within the specified range. Then, a Monte Carlo simulation of 1,000 iterations is executed to estimate the confidence interval.

The functions to calculate these indicators are `edible_jobs` and `volunteer_jobs` respectively. They share the same arguments, expect for `jobs` and `volunteers`, which is the value of $k$ in the previous equation. As usual, the first argument is the urban representation (`x`), the attribute of `x` defining the area used to grow plants is `are_col`, the confidence interval is defined in `interval`, and if `verbose` is set to `TRUE`, instead of the confidence interval, the function returns a vector of length 1,000 with all the results of the Monte Carlo simulation.

@@

edible_jobs(
  x,
  jobs = c(0.000163, 0.022),
  edible = NULL,
  area_col = "edible_area",
  interval = 0.95,
  verbose = FALSE
)

edible_volunteers(
  x,
  volunteers = c(0.00163, 0.22),
  edible = NULL,
  area_col = "edible_area",
  interval = 0.95,
  verbose = FALSE
)

@@

We propose an indicator to estimate green per capita at neighborhood level and at city level, including public and private gardens to account for environmental justice [@Kabisch2014c]. At the neighborhood level a ratio between the most and least green neighborhoods is calculated. Moreover, since wealthier areas tend to have more private gardens [@Farahani2018], these can be included in the account of green per capita to not underestimate green per capita in those neighbourhoods.

The function to calculate this indicator is `green_capita`. Along with the urban representation (`x`). As in other functions, the `green_categories` argument in the function is a list of the categories to be considered as green areas. To calculate the green per capita in the overall city, the user must provide the number of inhabitants in `inhabitants`. There are two options to calculate green per capita at a neighbourhood level. The urban representation can contain two variables indicating the neighbourhoods and the inhabitants (specified in `name_col` and `inh_col` arguments respectively). Or the user can provide a GIS layer with the neighbourhoods' boundaries and their attributes.

@@

green_capita(
  x,
  green_categories = NULL,
  inhabitants = NULL,
  neighbourhoods = NULL,
  name_col = NULL,
  inh_col = NULL,
  private = FALSE,
  verbose = FALSE,
  min_inh = 0
)

@@

Although many authors stated that the main goal of urban agriculture is not to produce food [@Saumel2019; @Tornaghi2012c], food production is undoubtedly an important component of urban gardens [@Furness2018; @Steel2008] and the most frequent output modeled at a city scale [@Grafius2020; @Grewal2012]. The food production is measured in terms of productivity:

$$
Food \; production = \sum_{i=1}^{N} \times y_k \times a_i
$$

where $y_k$ is the yield (in $kg/m^2$) of the category $k$ of urban garden; and $a_i$ is the area of urban garden $i$ in $m^2$. By default, the value of $y$ is randomized using a random uniform distribution within the range defined by `food1` and `food2` values in `city_functions`, which are the minimum and maximum yield values found in literature for each category of urban garden. The function computes a Monte Carlo simulation of 1000 iterations to calculate the confidence interval.

The function that calculates the food production is `food_production`. It takes the urban representation (`x`) as the first argument. If the second argument `edible_df` is `NULL`, the function uses the values from `city_functions` as specified above. Otherwise, the user can provide its own values as a `data.frame` with three columns:

- `functions`: specifying the category of urban agriculture, it should match the categories from `x`.
- `food1` and `food2` specifies the range of the random uniform distribution to randomize yield.

The argument `area_col` points to the variable of `x` that determines the area dedicated to grow plants in each urban garden. If `NULL`, the total area of each element is used instead. The number passed to `interval` defines which confidence intervals of the food production must be returned by the function. However, if `verbose` is set to `TRUE`, the function returns a vector of length 1,000 with the results of each iteration of the Monte Carlo simulation.

@@

food_production(
  x,
  edible_df = NULL,
  area_col = "edible_area",
  interval = 0.95,
  verbose = FALSE
)

@@

tibble(
  `Urban agriculture solutions` = city_functions$functions[1:5],
  Location = c("Private gardens", rep("Plots on ground", 2), rep("Rooftops", 2)),
  Purpose = c("Private", rep("Community", 2), rep("Commercial", 2))
) |> 
  kable(caption = "Elements created in new scenarios")

@@

The function to create a new scenario is called `set_scenario`. It requires many arguments but most of them have default values to facilitate its use. The function needs the urban representation (`x`). Then three arguments (`pGardens`, `pVacant`, `pRooftop`) control the proportion of new elements that must be created. The next three arguments (`edible_area_*`) control the proportion of the area of the new elements that is dedicated to grow plants (`edible_area`). The `edible_area` of each new elements is randomized within the range provided in the arguments. The next trio of arguments (`min_area_*`) specify the minimal area required to create new elements. If there are not enough elements largen than `min_area_*` to fulfill the first arguments, a message is displayed to inform the user (unless `quiet` argument is set to `TRUE`). Another three arguments (`*_from`) control which elements can be converted from the urban representation to create new urban agriculture solutions. The argument `pcommercial` controls the percentage of plots on ground and rooftop that should have commercial purposes instead of community. This does not affect private gardens since they are assumed to be for personal use. Finally, `area_field` specifies which attribute of `x` must be used as the area of the elements. By default, it is am attribute called `flat_area` that measures the area with an slope lower than 5º (in `city_example`).

@@

set_scenario(
  x,
  pGardens = 1,
  pVacant = 1,
  pRooftop = 1,
  edible_area_garden = c(0.02, 0.3),
  edible_area_vacant = c(0.52, 0.75),
  edible_area_rooftop = c(0.6, 0.62),
  min_area_garden = 10,
  min_area_vacant = 100,
  min_area_rooftop = 100,
  private_gardens_from = "Normal garden",
  vacant_from = "Vacant",
  rooftop_from = "Rooftop",
  pCommercial = 0,
  area_field = "flat_area",
  quiet = FALSE
)

@@

The `ediblecity` package is compatible with versions of R higher than 2.10. Although, it was created using version 4.2.1. `r ref("Code snippet", "install")` shows how to install the last development version, available in github.

@@

# install.packages("devtools") # if not yet installed
devtools::install_github("icra/ediblecity")

@@

To use the `ediblecity` package in R, it is necessary to install it first. `r ref("Code snippet", "install")` shows how to install the last development version.

```{r, eval=FALSE}

# install.packages("devtools") # if not yet installed
devtools::install_github("icra/ediblecity")

@@

Since the scenarios and the indicators have some stochastic parameters, the ideal procedure would be to integrate the creation of the scenarios and the estimation of indicators in a Monte Carlo simulation to get the confidence intervals for each combination of scenario and indicator. However, we rather keep things simple to better illustrate how to use the functions provided by the package.

@@

counting <- function(x, .name){
  x |> 
    st_drop_geometry() |> 
    filter(Function %in% city_functions$functions[1:5]) |> 
    group_by(Function) |> 
    summarize(!!sym(paste(.name, "n")) := n(),
              !!sym(paste(.name, "area")) := sum(area))
}

>>>>

---
title: "ediblecity: an R package to model and estimate the benefits of urban agriculture"
author:
- Josep Pueyo-Ros \textsuperscript{1,2,*}
- Joaquim Comas \textsuperscript{1,3}
- Lluís Corominas \textsuperscript{1,2}
output: 
  pdf_document:
    number_sections: true
    keep_tex: true

bibliography: library.bib
---

@@

Urban agriculture is gaining attraction to become one of the pillars of the urban ecological transition and to increase food security in an urbanized planet. However, there is a lack of systematic quantification of the benefits provided by urban agriculture solutions. In this paper, we present an R package to estimate several indicators related to benefits of urban agriculture. The goal is to provide a tool for researchers and practitioners interested in the impacts of urban agriculture. The ediblecity package provides functions to calculate 8 indicators: urban heat island, runoff prevention, green areas accessibility, NO~2~ sequestration, jobs created in commercial gardens, volunteers involved in community gardens, green per capita and, finally, food production. Moreover, the package also provides a function to generate scenarios with different implementations of urban agriculture. We illustrate the use of the package by comparing three scenarios in a neighborhood of Girona (Spain), which is included in the package as an example dataset. There, we compare scenarios with an increasing amount of urban agriculture solutions. The `ediblecity` package is open-source software. This allows other R developers to contribute to the package providing new functionalities or improving the existing ones.

@@

Therefore, decision-makers, who are responsible of leading the urban transitions to more sustainable and resilient cities, are orphan of evidence in how to implement urban agriculture to maximize its impact on sustainability. Yet, several studies provided some insights that can guide the implementation of urban agriculture. For instance, some models explore the rainwater harvesting potential of urban agriculture [@Gittleman2017; @Lupia2017]. Broader, @GomezVillarino2021 developed guidelines to maximize ecosystem services through urban agriculture by applying adaptive design and providing a battery of indicators. And, as expected, many models have been developed to quantify food production and food security provided by urban agriculture by simulating a myriad of scenarios with different types of urban agriculture virtually implemented such as rooftop gardens, community gardens or private citizen-led gardens [@Grafius2020; @Grewal2012; @Hsieh2017; @MacRae2010]. Yet, a transferable model (applicable to any city) to assess simultaneously several environmental and social benefits is lacking.

Hence, our goal is to provide a tool to estimate those multiple benefits of urban agriculture that help decision-makers to strategically implement urban agriculture solutions. We developed the estimations for eight indicators measuring urban agriculture benefits and a functionality to create scenarios of urban agriculture based on the proportion of elements to be transformed to urban agriculture and on which elements will be transformed. Likewise, we packed all those functionalities in an R package called `ediblecity`. In the methods, we present the interface of the package and then we illustrate the usefulness by applying the model in a neighborhood of Girona (Spain).

@@

## Implementation: the model under the package

@@

The `ediblecity` package provides 8 functions to estimate 8 different indicators and a function to generate scenarios. It takes as a basis a spatial representation of a city (or a part of a city) as a GIS layer. This representation must have one attribute indicating the land uses of the city, such as gardens, streets, rooftops, etc. Some indicators require specific information to be included in the representation. These is explained in each indicator's section.

```{r}
color_names <- sort(unique(city_example$land_use))
colors <- c("Brown", "Lightgreen", "chartreuse4", "Grey20", 
            "Black", "chocolate2", "Yellow", "Lightblue",
            "Red", "Grey50", "Darkgreen", "Yellow"
            )
names(colors) <- color_names

@@

city_example |> 
  ggplot()+
  geom_sf(aes(fill=land_use), color="white", linewidth = 0.1)+
  scale_fill_manual(values = colors)+
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        axis.text = element_text(size=6))

@@

tribble(
  ~Column, ~Description,
  "land_use", "A category representing the urban elements",
  "land_use_verbose", "A more detailed category for the element, for example, if it is residential",
  "floors", "The number of floors of the element, 0 for non-built",
  "area", "The surface of the element",
  "flat_area", "The surface of the element that is flat (slope < 5º)",
  "edible_area", "The surface that is used to grow edible plants. Only applicable to urban agriculture solutions"
) |> 
  kable(caption = "Structure of the urban representation example")

@@

Along with the example for an urban representation, the `ediblecity` package also includes a `data.frame` with the default attributes of each green typology in the urban representation used to estimate the indicators (Table 2). This includes types of urban agriculture along with other types of green infrastructure. However, the user can provide their own attributes to estimate any indicator. In `city_land_uses` there are other columns not shown in table 2, they are logical variables (i.e. TRUE/FALSE) used internally by the package to select urban agriculture elements.

`pGreen` is the proportion of green of the urban element. In urban agriculture solutions, this is overriden by the attribute `edible_area`. The following attributes come in pairs (min, max) to consider uncertainty in the estimations. The functions use a random value within the range provided by the pair of values for each element in the city. `no2_seq` is the capacity of the element to capture NO~2~ in gr/s. `food` is the food productivity in kg/m^2^ and `CN` is their curve number, used to calculate infiltration rates. The details are provided in following sections.

@@

city_land_uses |>
  select(-starts_with("...")) |> 
  select(-edible, -public, -jobs, -volunteers, -location) |>
  select(-starts_with("water")) |> 
  kable(caption = "General attributes of the elements of urban green used to estimate the indicators")
```
### Indicators estimated

@@

The urban heat island is a measure of how urban agriculture can contribute to climate change adaptation, specifically to adapt cities to the increasing heat waves [@Langemeyer2021]. The indicator uses the equation developed by @Theeuwes2017, which was validated in 14 cities. It calculates the difference in air temperature between the urban street canyon and the rural environment. To calculate this indicator, the user must provide a raster representing the sky view factor (SVF), which describes the proportion of the unobstructed hemisphere above a certain location. SAGA, a collection of open-source algorithms for geocomputation, provides an algorithm to calculate the SVF [@Conrad2008].

@@

The indicator to estimate the urban heat island is implemented by the package under the function `UHI`. The user must provide the urban representation (`x`) and the raster with SVF values (`SVF`). The `green_df` argument is a `data.frame` with the proportion of green of each `land_use` in the urban representation ($Fveg$ in the equation). All the meteorological arguments are provided by default, based on the example provided (Mediterranean climate). However, the user can override them to provide values of their city of interest.

@@

UHI(
  x,                         # Urban representation
  SVF,                       # Raster with the sky view factor
  green_df = NULL,           # Fveg in the UHI equation
  Qql = 6.11,                # Daily average global radiation
  Cair = 1007,               # Air heat capacity
  Pair = 1.14,               # Air density
  Tmax = 30.8,               # Maximum daily average temperature
  Tmin = 20,                 # Minimum daily average temperature
  windspeed = 2.77,          # Average wind speed
  return_raster = FALSE,     # Should function return a raster or numeric values?
  verbose = FALSE            # Should the function return all the values or a summary?
)

@@

where $Q$ is the runoff in $mm$; $P$ is the rainfall volume in mm; $I_a$ is the initial abstraction (all losses before runoff begins); and $S$ is the potential soil moisture retention, which is a function of the curve number (which is determined by hydrologic soil group, see @Cronshey1985 for more details on the method). The SCS generalizes $I_a$ as $0.2S$, we modified this generalization to include the rainwater harvested:

@@

The `runoff_prev` function estimates the runoff ($Q$ in the equation) as well as the total rainfall in `x` and the total rainwater harvested in cubic meters. Along with the urban representation (`x`), the user must provide a `data.frame` (`runoff_df` argument) with four variables: land use, minimum and maximum curve numbers and a logical variable indicating if the land use (a urban garden, a building,...) has potential to harvest rainwater. The argument `rain` allows to set the rain event (in mm), which must be defined by the user. The curve number of each element is randomized within this range provided in `runoff_df`. If `runoff_df` is not provided, `city_land_uses` is used instead. Following, `floors_field` is the name of attribute in `x` that specifies the number of floors of each element; `harvest_dist` (in m) is the maximum distance to consider that a building is adjacent to the element; and `tank_size` is a range for the volume of the rainwater tank, proportional to the surface of the element (in l/m^2^). The volume of each tank in the city is randomized within this range and multiplied by the element where they are located. All randomizations follow a random uniform distribution within the correspondent range.

@@

runoff_prev(
  x,                          # Urban representation
  runoff_df = NULL,           # A data.frame with land use values
  rain = 85,                  # Rain event in mm.
  floors_field = "floors",    # Variable in `x` containing the number of floors of each element
  harvest_dist = 10,          # Maximum distance to consider for rainwater harvesting surfaces
  tank_size = c(0, 45)        # Range for the size of tanks in m3
)
```

@@

This indicator calculates the distance from each home to the closest public green area and return a summary of statistics (min, 25%, 50%, mean, 75% and max). It includes the possibility to exclude areas smaller than a threshold. The function to calculate these distances is `green_distance` (code snippet 3). It requires, as usual, the urban representation (`x`) and a vector with all the categories considered public green areas. If it is not provided, the function uses the categories from `city_land_uses` where the attribute `public` is `TRUE`. The argument `residence_col` indicates the variable of the urban representation (`x`) that must be used to identify the residences. Subsequently, `residences` indicates which categories of the ones contained by the variable passed to `residence_col` must be considered.  The `min_area` argument can be used to exclude smaller areas than the value passed to the function (the threshold mentioned above).

@@

Finally, if `verbose` argument is set to `TRUE`, a vector with all distances from residences to green areas (larger than `min_area`) is returned.

@@

green_distance(
  x,                                     # The urban representation
  green_cat = NULL,                      # The land uses considered urban green
  residence_col = "land_use_verbose",    # The variable than contains information of what is a residence
  residences = "Residence",              # The categories of former variable considered residences
  min_area = 5000,                       # Smaller areas are not included in the calculations
  percent_out = FALSE,                   # Should the function return a percentage of residences out
  max_dist = 300,                        # Which is the distaince to consider that a residence is out?
  verbose = FALSE                        # Should return a summary of distance or all values?
)

@@

$$
NO_2 seq = \sum_{i=1}^{N}\frac{{a_i \times cap_i}}{1000}
$$

where $a_i$ is the area (in m^2^) of the element $i$; and $cap_i$ is the capacity of element $i$ to sequester NO~2~ (in $\mu$g·s^-1·m^-2^).

The function to estimate the NO~2~ sequestered is `no2_seq` (code snippet 4). It has only two arguments; the urban representation (`x`) and a `data.frame` with four columns:

- `land_uses`: Column with the land use.
- `no2_seq1`: The low range of NO2 sequestration of each function (in $\mu$g·s^-1/m-^2^).
- `no2_seq2`: The high range of NO2 sequestration of each function (in $\mu$g·s^-1/m-^2^).
- `pGreen`: The proportion of green surface in each function.

When the argument is `NULL` (default), the function uses the `city_land_uses` dataset provided with the package that contains NO~2~ sequester capacity of different types of urban green.

@@

no2_seq(x,                 # The urban representation
        green_df = NULL)   # The data.frame with the NO2 sequestration capacity of each land use.

@@

When the urban agriculture solutions are community solutions they need volunteers to be involved. On the other hand, when they are for commercial purposes, they are supposed to create jobs. Therefore, two indicators are proposed to account for the hours of work in the ECS. One indicator relates to volunteers time and the other one relates to time of new workers (jobs created).  Both indicators use the same equation:

@@

where $a_i$ is the area in m^2^ used to grow plants (`edible_area`) in the element $i$; and $k$ is the number of jobs or volunteers by m^2^. In both functions, $k$ is sampled from a random uniform distribution within the specified range. Then, a Monte Carlo simulation of 1,000 iterations is executed to estimate the confidence interval. The default values were based on two sources, an empirical study [@CoDyre2015] and the [FoodMetres project](http://www.foodmetres-kp.eu).

The functions to calculate these indicators are `edible_jobs` and `edible_volunteers` respectively (code snippet 5). They share the same arguments, expect for `jobs` and `volunteers`, which is the value of $k$ in the previous equation. As usual, the first argument is the urban representation (`x`), `edible` is the land uses in `x` that are urban agriculture solutions (if `NULL`, `city_land_uses` is used as default), the attribute of `x` defining the area used to grow plants is `area_col`, the confidence interval is defined in `interval`, and if `verbose` is set to `TRUE`, instead of the confidence interval, the function returns a vector of length 1,000 with all the results of the Monte Carlo simulation.

@@

edible_jobs(
  x,                              # The urban representation
  jobs = c(0.000163, 0.022),      # The k parameter
  edible = NULL,                  # The land uses that are urban agriculture
  area_col = "edible_area",       # The variable containing the surface dedicated to grow plants
  interval = 0.95,                # The confidence interval returned by the function
  verbose = FALSE                 # Should return all values or a summary?
)

edible_volunteers(
  x,                              # The urban representation
  volunteers = c(0.00163, 0.22),  # The k parameter
  edible = NULL,                  # The land uses that are urban agriculture
  area_col = "edible_area",       # The variable containing the surface dedicated to grow plants
  interval = 0.95,                # The confidence interval returned by the function
  verbose = FALSE                 # Should return all values or a summary? 
)

@@

We propose an indicator to estimate green per capita at neighborhood level and at city level, including public and private gardens to account for environmental justice [@Kabisch2014c]. At the neighborhood level a ratio between the most and least green neighborhoods is calculated, thus higher values meaning a major difference between neighborhoods and less spatial justice. Moreover, since wealthier areas tend to have more private gardens [@Farahani2018], these can be included in the account of green per capita to not underestimate green per capita in those neighborhoods.

The function to calculate this indicator is `green_capita` (code snippet 6). Along with the urban representation (`x`). As in other functions, the `green_categories` argument in the function is a list of the categories to be considered as green areas. To calculate the green per capita in the overall city, the user must provide the number of inhabitants in `inhabitants`. At a city level, the function returns the average amount of green per capita in the city (in m^2^/inhabitant). Likewise there are two options to calculate green per capita at a neighbourhood level. The urban representation can contain two variables indicating the name of the neighbourhood and the inhabitants (specified in `name_col` and `inh_col` arguments respectively). Or the user can provide a GIS layer with the neighbourhoods' boundaries and their attributes.

@@

green_capita(
  x,                        # The urban representation
  green_categories = NULL,  # The categories considered as urban green
  inhabitants = NULL,       # The number of inhabitants in the city
  neighbourhoods = NULL,    # The spatial representation of neighborhoods
  name_col = NULL,          # The variable that contains de name of the neighborhoods
  inh_col = NULL,           # The variable that contains the inhabitants of the neighborhoods
  private = FALSE,          # Should private green be considered?
  verbose = FALSE,          # Should return all the information or just the ratio?
  min_inh = 0               # Neighborhoods with less hanbitants are excluded
)

@@

Although many authors stated that the main goal of urban agriculture is not to produce food [@Saumel2019; @Tornaghi2012c], food production is undoubtedly an important component of urban gardens [@Furness2018; @Steel2008] and the most frequent output modeled at a city scale [@Grafius2020; @Grewal2012]. The food production is measured in terms of productivity (kg/m^2^):

$$
Food \; production = \sum_{i=1}^{N}y_k a_i
$$

where $y_k$ is the yield (in $kg/m^2$) of the category $k$ of urban garden; and $a_i$ is the area of urban garden $i$ in $m^2$. By default, the value of $y$ is randomized using a random uniform distribution within the range defined by `food1` and `food2` values in `city_land_uses`, which are the minimum and maximum yield values found in literature for each category of urban garden. The function computes a Monte Carlo simulation of 1000 iterations to calculate the confidence interval.

The function that calculates the food production is `food_production` (code snippet 7). It takes the urban representation (`x`) as the first argument. If the second argument `edible_df` is `NULL`, the function uses the values from `city_land_uses` as specified above. Otherwise, the user can provide its own values as a `data.frame` with three columns:

- `land_uses`: specifying the category of urban agriculture, it should match the categories from `x`.
- `food1` and `food2` specifies the range of the random uniform distribution to randomize yield.

The argument `area_col` points to the variable of `x` that determines the area dedicated to grow plants in each urban garden. If `NULL`, the total area of each element is used instead. The number passed to `interval` defines which confidence interval of the food production must be returned by the function. However, if `verbose` is set to `TRUE`, the function returns a vector of length 1,000 with the results of each iteration of the Monte Carlo simulation.

@@

food_production(
  x,                        # The urban representation
  edible_df = NULL,         # Dataframe containing information on yields
  area_col = "edible_area", # Variable of x with surface
  interval = 0.95,          # Confidence interval returned by the function
  verbose = FALSE           # Should return all the values or just the confidence interval?
)

@@

tibble(
  `Urban agriculture solutions` = city_land_uses$land_uses[1:5],
  Location = c("Private gardens", rep("Plots on ground", 2), rep("Rooftops", 2)),
  Purpose = c("Private", rep("Community", 2), rep("Commercial", 2))
) |> 
  kable(caption = "Elements created in new scenarios")

@@

The function to create a new scenario is called `set_scenario` (code snippet 8). It requires many arguments but most of them have default values to facilitate its use. The function needs the urban representation (`x`). Then three arguments (`pGardens`, `pVacant`, `pRooftop`) control the proportion of new elements that must be created. The next three arguments (`edible_area_*`) control the proportion of the area of the new elements that is dedicated to grow plants (`edible_area`). The `edible_area` of each new elements is randomized within the range provided in the arguments. The next trio of arguments (`min_area_*`) specify the minimal area required to create new elements. If there are not enough elements largen than `min_area_*` to fulfill the first arguments, a message is displayed to inform the user (unless `quiet` argument is set to `TRUE`). Another three arguments (`*_from`) control which elements can be converted from the urban representation to create new urban agriculture solutions. The argument `pcommercial` controls the percentage of plots on ground and rooftop that should have commercial purposes instead of community. This does not affect private gardens since they are assumed to be for personal use. Finally, `area_field` specifies which attribute of `x` must be used as the area of the elements. By default, it is am attribute called `flat_area` that measures the area with an slope lower than 5º (in `city_example`).

@@

set_scenario(
  x,                                       # The urban representation
  pGardens = 1,                            # Proportion of private edible gardens
  pVacant = 1,                             # Proportion of vacant plots converted to gardens on ground
  pRooftop = 1,                            # Proportion of rooftops converted to rooftop gardens
  edible_area_garden = c(0.02, 0.3),       # Proportion of surface dedicated to grow plants in private gardens
  edible_area_vacant = c(0.52, 0.75),      # And in gardens on ground
  edible_area_rooftop = c(0.6, 0.62),      # And in rooftop gardens
  min_area_garden = 10,                    # Exclude smaller private gardens
  min_area_vacant = 100,                   # Exclude smaller vacant plots
  min_area_rooftop = 100,                  # Exclude smaller rooftops
  private_gardens_from = "Normal garden",  # Land uses to be converted to private edible gardens
  vacant_from = "Vacant",                  # Land uses to be converted to gardens on ground
  rooftop_from = "Rooftop",                # Land uses to be converted to rooftop gardens
  pCommercial = 0,                         # Proportion of commercial gardens vs community gardens
  area_field = "flat_area",                # Variable of x containing available surface in each element
  quiet = FALSE                            # Should the function raise warnings?
)

@@

The `ediblecity` package is compatible with versions of R higher than 2.10. Although, it was created using version 4.2.1. `r ref("Code snippet", "install")` shows how to install the last development version, available in [r-universe](https://r-universe.dev/search/).

@@

# install.packages("devtools") # if not yet installed
install.packages("ediblecity", repos = "jospueyo.r-universe.dev")

@@

@@

@@

counting <- function(x, .name){
  x |> 
    st_drop_geometry() |> 
    filter(land_use %in% city_land_uses$land_uses[1:5]) |> 
    group_by(land_use) |> 
    summarize(!!sym(paste(.name, "n")) := n(),
              !!sym(paste(.name, "area")) := sum(area))
}

counting(scenarios[[1]], "s0") |> 
  full_join(counting(scenarios[[2]], "s1")) |> 
  full_join(counting(scenarios[[3]], "s2")) |> 
  mutate(across(where(is.numeric), ~if_else(is.na(.x), 0L, .x))) |> 
  relocate(land_use, ends_with("n"), ends_with("area")) |> 
  kable(caption = "Number of agriculture solutions in each scenario and their surfaces (in squared meters)")
```
